---
title: "Lab 3: Post Lab"
format: html
---

## Data Load
```{r}
library(tidyverse)
library(tidymodels)
library(here)

```

```{r}
#import data from this website

wine_quality <- readr::read_delim("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv",delim=';')
 
```


## Design a research question

Examine the data to come up with an interesting research question based on the wine data. It is a large collection of wine attributes and quality ratings to indicate the sommlier's rating of the wine.

change quality into binary data and fit different parameters to a logistic model 

## Explore the data
Create some exploratory tables or charts to help you better understand possible relationships


```{r}
head(wine_quality)

wine_df -> wine_quality %>% 
  select(pH, density) %>% 
  
  
```

## Use a Tidymodel framework to conduct your analysis.

## Visualize

Present your final results and upgrade this document's YAML to a professional level quality.
